---
title: "Homework Sheet 2 -- Wrangle, Summarize, Visualize -- Excercises"
date: 'Due: Wednesday, November 18 by 11:59 CET'
author: ""
output: 
  html_document:
    toc: true
    toc_depth: 2
    highlight: tango
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, error = F, message = F, warning = F)

```

```{r libraries, include=FALSE, message=FALSE, warning=FALSE}

# package for convenience functions (e.g. ggplot2, dplyr, etc.)
library(tidyverse)

# package for this course
library(aida)

# use the aida-theme for plotting
theme_set(theme_aida())

# global color scheme / non-optimized
project_colors = c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#000000")

# setting theme colors globally
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = project_colors)
}
scale_fill_discrete <- function(...) {
   scale_fill_manual(..., values = project_colors)
} 

# nicer global knitr options
knitr::opts_chunk$set(warning = FALSE, message = FALSE, 
                      cache = TRUE, fig.align = 'center')


```

# Instructions

* If you need help, take a look at the suggested readings in the lecture, make use of the [cheat sheets](https://rstudio.com/resources/cheatsheets/) and the help possibility in R
* Make sure you have R and RStudio installed. If you are an advanced user and aren't using RStudio, make sure you have everything installed in order to 'knit' the HTML output.
* Create an Rmd-file with your group number (equivalent to StudIP group) in the 'author' heading and answer the following questions.
* When all answers are ready, 'Knit' the document to produce a HTML file.
* Create a ZIP archive called "IDA_HW02-Group-XYZ.zip" (where 'XYZ' is *your* group number) containing:
   * an R Markdown file "IDA_HW02-Group-XYZ.Rmd"
   * a knitted HTML document "IDA_HW02-Group-XYZ.html"
   * include the data files 'WHO.csv' and 'BLJM-data-2020.csv' as well
   * any other files necessary to compile your Rmarkdown to HTML (pictures etc.)
* Upload the ZIP archive on Stud.IP in your group folder before the deadline (see above). You may upload as many times as you like before the deadline, only your final submission will count.

# <span style = "color:firebrick">Exercise 1:</span> Explore your own BLJM data [15 points]

Make sure you obtained the file `BLJM-data-2020.csv` from StudIP. It contains data from the "Bio-Logic Jazz-Metal" experiment, which is also coverd in the web-book. But this data is what *you* have produced by taking part in the voluntary web-experiment. So, let's go analyze **YOUR** preferences!

## Ex 1.a Global summary statistics [2 points]

Load the file using the code below (the CSV file should be in the same directory as your Rmd file).

```{r }
data_BLJM <- read_csv("BLJM-data-2020.csv")
```

Use R to display the following information:

1. the number of rows in the data set
2. the number of different entries in the column `submission_id` (many different solutions possible, try a combination of `group_by`, `summarize` and `nrow` if you like)
3. the mean `timeSpent`
4. the mean `age` (*Hint:* use the option `mean(.., na.rm=T)` to remove entries with `NA`)


## Ex 1.b  Grouped summary statistics [2 points]

Use piping to perform the following steps on `data_BLJM`: 

1. throw out all rows where the column `gender` has an entry of `NA` (*Hint*: use function `is.na` and Boolean negation); 
2. group by `gender`
3. obtain the mean and (unbiased) standard deviation for `age` (exluding `NA`s in both functions (see Ex 1.a))


## Ex 1.c Preprocessing [2 points]

We will use the code from [Appendix D of the web-book](https://michael-franke.github.io/intro-data-analysis/app-93-data-sets-BLJM.html) to process the data for more convenient analysis. Insert short but insightful comments in the code below:

```{r}
data_BLJM_processed <- 
  data_BLJM %>% 
  # ... your exlanation here ...
  mutate(
    condition = str_c(str_sub(option2, 1, 1), str_sub(option1, 1, 1))
  ) %>% 
  # ... your exlanation here ...
  select(submission_id, condition, response)

```


## Ex 1.d Obtaining counts for all three choices [2 points]

Produce a tibble with 8 rows which contains the following tidy data for later processing: 

1. a character column `JM` which contains entries for Jazz vs Metal
2. a character column `LB` which contains entries for Biology vs Logic
3. a character column `BM` which contains entries for Beach vs Mountain
4. a column that contains the number of times the triple of choices in each row occurred in the data

(*Hint*: You can recycle code from the web-book, but you need to find it (and possibly tweak it).)


## Ex 1.e Obtaining counts for music+subject choices [2 points]

Produce a similar table as for the previous exercise but only for the `JM` and `LB` columns.


## Ex 1.f Plot the music+subject choices [2 points]

Create a plot for the current data set similar to that of Figure 5.1 from the web-book. Comment your code exetensivly, explaining each relevant step as if you were trying to teach somebody R and ggplot. 

(*Hint*: you can recycle code from the web-book, but you need to find it.)


## Ex 1.g Interpret the plot [3 points]

Argue, if you can, that the lecturer's conjecture is true: students who like metal also prefer logic, and vice versa. Or, if you must, argue that the the data does not provide (visual) evidence for the lecturer's conjecture. (Make your answer short; at most four concise sentences, please!)


# <span style = "color:firebrick">Exercise 2:</span> Musings about modes [5 points]

Suppose we have a vector of metric data points, stemming from measurements of weight of around ten thousand seagulls. The weight has been determined up to the [yoctogram level](https://en.wikipedia.org/wiki/Orders_of_magnitude_(mass)). (Just imagine that this was actually possible!) Why would it probably not make much good sense to consider the mode of this vector as an indicator of central tendency of these measurements?


# <span style = "color:firebrick">Exercise 3:</span> Exploring two-dimensional metric measurements [15 points]

We are now going to practice plotting and analyzing metric data from more than one variable. To do this, we first load the WHO data set with the code below. 

```{r}
data_WHO <- read_csv('WHO.csv')
```

The WHO data set has a number of health-related measurements for most of the world's nations, grouped by region. It is useful for you if you try to explore the data set on your own, just to see what is there. (*Hint:* try `View(data_WHO` in RStudio).


## Ex 3.a Calculate a correlation score [2 points]

Use R's built-in function `cor` to compute the Bravais-Pearson correlation between the variables `FertilityRate` and `LifeExpectancy`. Notice that you must first remove all entries in either column with missing data.


## Ex 3.b Interpret a correlation score [3 points]

Interpret this numerical result. Your answer be something like the following (but hopefully much better): "There appears to be a positive correlation between $X$ and $Y$ suggesting that as $X$ increases $Y$ will decrease. This is proof that $X$ is a causally related to $Y$, in fact that $Y$ is the cause of $X$."


## Ex 3.c Compute correlation scores for each region [2 points]

Use R's built-in function `cor` to compute the Bravais-Pearson correlation between the variables `FertilityRate` and `LifeExpectancy` seperately for each region. Use `group_by` and remember to remove missing values. Store the resulting tibble in a variable called `correlation_WHO` because we will use it later on.


## Ex 3.d Make a scatter plot [3 points]

Create a scatterplot (using `geom_point`) showing `FertilityRate` on the $x$-axis and `LifeExpectancy` on the $y$-axis. Add a linear regression line, using `geom_smooth`. (*Hint*: See Section 6.3.3 of the web-book). For good sports, check the documentation of `geom_smooth` to find an option that suppresses the gray area around the regression line (which is a confidence interval; a concept we will learn about later in the course). Finally, adjust the axis labels, so as not to show CamelCaseLabels but "Nice labels with spaces".


## Ex 3.e Make a scatter plot facetted by region [2 points]

Use `facet_wrap` to make six nicely arranged scatter plots, one for each level of `Region`, of `FertilityRate` against `LifeExpectancy` (as before: with linear regression lines w/o confidence intervals and nice axis labels). Make sure that each facet has its own range of values shown for the $x$- and $y$-axis, using the `scales` option of `facet_wrap`.


## Ex 3.f Add labels to each facet [3 points]

(*Caveat:* This exercise might be more challenging than the others.)

Add the information stored in `correlation_WHO` from Ex. 3.f to each facet using `geom_text`. The following code will help you.

```{r, echo=TRUE, eval=FALSE}
correlation_WHO <- 
  correlation_WHO %>% 
  mutate(
    x = c(6, 3.5, 6, 3.5, 5, 4),
    y = c(70, 77, 70, 75, 72, 78), 
    correlation = round(correlation, 2)
  )
```

